---
title: "campaign_food_spending"
author: "Dylan Manfre"
date: "2023-08-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#libraries
library(tidyverse)
library(tidycensus)
library(janitor)
library(lubridate)
#install.packages("ipumsr")
library(ipumsr)
```

*** Early notes from Aug. 23 ***

- Years filtered for: 2011 to today
- States filtered fsor: DC, MD, VA
- Search terms I used in the FEC.gov site.
  - MEALS, CATERING, FOOD, FOOD/BEVERAGE, FOOD AND BEVERAGE EXPENSES, CATERING EVENTS, CATERING FOOD AND BEVERAGES, DRINKS/FOOD (PRODUCED 213,860 RESULTS)
    -LINK: https://www.fec.gov/data/disbursements/?data_type=processed&two_year_transaction_period=2012&two_year_transaction_period=2014&two_year_transaction_period=2016&two_year_transaction_period=2018&two_year_transaction_period=2020&two_year_transaction_period=2022&two_year_transaction_period=2024&min_date=01%2F01%2F2011&max_date=12%2F31%2F2021&recipient_state=DC&recipient_state=MD&recipient_state=VA&disbursement_description=FOOD&disbursement_description=MEALS&disbursement_description=CATERING&disbursement_description=FOOD%2FBEVERAGE&disbursement_description=FOOD+AND+BEVERAGE+EXPENSES&disbursement_description=CATERING+EVENTS&disbursement_description=CATERING+FOOD+AND+BEVERAGES&disbursement_description=CATERING%2FEVENTS&disbursement_description=DRINKS%2FFOOD
  
  
  
  - Use single words. not phrases
  

  

- We want to identify the purpose category as "Fundraising" or another term and make sure it doesn't say "travel" so we exclude those unnecessary transactions.



```{r}

# loading in the campaign_food_data.csv
campaign_food_data <- read_csv("data/campaign_food_data.csv")
head(campaign_food_data)

```


** I put in 7 filters to the FEC site with the above link. Shouldn't then there only be 7 categories that show up? **

```{r}
campaign_food_data %>% 
group_by(disbursement_description) %>% 
count() %>% 
arrange(desc(n))

campaign_food_data %>% 
group_by(recipient_name, recipient_state) %>% 
summarize(total_spent = sum(disbursement_amount)) %>% 
arrange(desc(total_spent))



```

** Need to use Open Refine to standardize some of the disbursement_descriptions
List of ones to standardist: FOOD AND BEVERAGE, FOOD & BEVERAGE, FOOD/BEVERAGES, FOOD / BEVERAGES, FOOD AND MEALS, FOOD & BEV, FOOD AND BEVERAGE EXPENSE

** RESTURANT MEALS IS A GOOD ONE TO HONE IN ON **







