---
title: "bls_work"
author: "Dylan Manfre"
date: "2023-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Establishing my steps

```{r}

# What are my steps here and can I create a loop from it?
# import the dataset wbb_rosters_22_23
# establish the power_five list power_five <- c("SEC", "Big Ten", "Pac-12", "Big 12", "ACC")

#Try to set this up so that the code is underneith a sentence that I would write in the story. 


```

*** Genearl notes *****

Thinking of making this story more of a listical piece where I provide a graphic that answers questions I have when writing the code and analyzing the data. Each section of the story would be an answered question about the women's basketball data. And all of them will be relating back to Maryland or Power 5 at large.


General questions
- Do I start national and work down to a more local level?
- Do I include schools that are moving into the Big Ten in the latest 23-24 data?
- How do I work in the census data to get per capita numbers (look back at the textbook and previous labs)


Steps
- Generate the numbers per capita for the state and then the local towns
- Get state populations by using Census API 
- Begin writing the questions for the story
- Start national and then work down




Story questions to be analyzed by the data (In no particular order)
- Where to women's basketball players in Maryland come from
- Where do Big 10 players come from in Maryland?
- Which county or state produces the most women's basketball players?
- How many Power 5 players are there in Maryland?
- Which state has the most WBB players per capita?
- How many DI, II and III players are there. 


Story Structure (Start wide and work your way in)
- National numbers
- Maryland state numbers
- Basketball centric numbers: ie: Big 10, Division I etc.




Presentation ideas
- Vision is tha you click on a part of the map and then it zooms in on an area and you're able to or click on a filter like the first graphic in the Ramaswamy story



```{r}
#libraries
library(tidyverse)
library(tidycensus)
library(janitor)
library(lubridate)
#install.packages("ipumsr")
library(ipumsr)

```

```{r}

#Loading in the data from 22-23 season

wbb_rosters_22_23 <- read_csv("data/wbb_rosters_2022_23.csv")
head(wbb_rosters_22_23)

#Loading in 23-24 data when I have it


```

*** Organized section ****


Working Lede: With women's basketball season approaching, let's see where players come form within the US and abroad. While Maryland lost a plethora of talent, including three of its starters, the Terps have a good crop of experience and newcomers to work with.


Question: What state has the most WBB players
Answer: PA has the most with 915
Sentence: There are over 1,000 players who hail from overseas (say some of the best international names), but Pennsylvania, New York and Texas dominate the domestic states with 915, 900 and 992 players respectively.
```{r}

state_counts <- wbb_rosters_22_23 %>% 
group_by(state_clean) %>%
count() %>% 
arrange(desc(n))
#1073 NAs

international_players <- wbb_rosters_22_23 %>% 
filter(is.na(state_clean))

```

Question: How many WBB players are from Maryland? How many of those players attend power 5 schools.?
Answer: 350 are from Maryland and 24 of them attend Power 5 schools
Sentence: Roughly 7% of all Maryland wbb players attend power five schools.
```{r}

power_five <- c("SEC", "Big Ten", "Pac-12", "Big 12", "ACC")

wbb_rosters_22_23 %>% 
filter (state_clean == "MD")
#produces 350 names

from_maryland_p5 <-  wbb_rosters_22_23 %>% 
filter(conference %in% power_five, state_clean == "MD")
#Produces 24 names

#Percentage of MD players at P5 institutions
md_p5_perc <-  (24 / 350)*100
  
```


Question: Which power 5 conference produces the most WBB players?
Answer: There are 875 power 5 wbb players and 196 of them are from the ACC, the mos of the fiv econferneces.
Sentence: It's pretty hard to land a scholarship offer. (Insert stat about college athletes)/ Getting a spot in one of the five conferences (soon to be 4?) is even harder. Only 875 women's basektball players attend Power 5 institutions. The ACC is most likely to take you. They have 196 players, but not much in front of the Big Ten with 194.
```{r}

wbb_rosters_22_23 %>%
filter(conference == power_five)

#What I want to do here is make a value that says power_five and have that be equal to "Big Ten", "Big 12", "ACC", "SEC', "Pac-12" so that when I filter I can say Filter to show me only players in the power_five conferences.

# Issue: These numbers seem too low. Im trying to view how many players there are in the power 5 conferneces
# Why doesn't every roster show up?

#Second attempt
# ChatGPT says to use in%in instead of %> %>% 

total_p5_players <- wbb_rosters_22_23 %>%
filter(conference %in% power_five)
#This produces 875 results which is much cleaner and seems to be more accurate. 

#Conference counts with percentages
conf_player_counts <- total_p5_players %>% 
group_by(conference) %>%
summarize(total_players = n()) %>% 
arrange(desc(total_players)) %>% 
mutate(perc = round(total_players/sum(total_players)*100,1))

```

#Question: How many players from Maryland does each power 5 confernece have?
#Answer: ACC and Big Ten each have 8 and there are no players from the Big 12
#Sentence: The lone Pac-12 player from Maryland will soon jump to the Big Ten as UCLA is set to join in Aug. 2024

Note: Decide what to do with conference realignment teams here.
```{r}

conference_md <- from_maryland_p5 %>% 
group_by(conference) %>% 
count() %>% 
rename(players_from_md = n)

#Remember to factor in conference realignment...

```

Question: Where do the Power 5 Players come from in Maryland specifically?
Answer: Baltimore has the most with 6 players on P5 tems. It also has the most players of any division.
Sentence:


Note: Decide which is more focused for the story
```{r}
# Power 5 players from Maryland
md_p5_hometown <- from_maryland_p5 %>% 
group_by(hometown_clean) %>% 
count() %>% 
arrange(desc(n)) %>% 
rename(players_from_md = n)

#All Players from Maryland
all_md_players_hometown <-  wbb_rosters_22_23 %>% 
filter(state_clean == "MD") %>% 
group_by(hometown_clean) %>% 
count() %>% 
arrange(desc(n)) %>% 
rename(players_from_md = n)
view(all_md_players_hometown)


```

Question: Show me a division breakdown of the players in the dataset that are from Maryland
Answer: There are only 133 D I players from Maryland and 140 who are in D III.
Sentence: 18% of players from Maryland play at P5 institutions. UMD has the most with 3 players but Arkansas is a surprise with 2 players since geographically it's further away. 


```{r}
md_division_breakdown <- wbb_rosters_22_23 %>% 
filter(state_clean == "MD") %>% 
group_by(division) %>% 
count() %>% 
arrange(desc(n)) %>% 
rename(players_from_md = n)

#Note: Putting here Percentage of D I players at Power 5 schools.
#There are 24 P5 players at Division I schools

md_d1_p5_perc = (24/133)*100
#18% of division I players from Maryland attend P5 schools.

from_maryland_p5 %>% 
group_by(team) %>% 
count() %>% 
arrange(desc(n))

```





*** Beginning to work on the section with Census Data ***

```{r}
#API key
Sys.setenv(CENSUS_API_KEY = "5cb2b9c628a1d1329c5bf4d36fe4435c6501efc8", overwrite = TRUE)

#2021 ACS
ACS_2021 <- load_variables(2021, "acs5")

#2022 ACS
ACS_2022 <- load_variables(2022, "acs1") #Figured that ACS 2022 would be best for this project because im only comparing 2022-2023 and 23-24

#Get this to show the population data of all the Maryland towns
md_2022 <- get_acs(geography = "place",
                variables = c(population = "B01003_001"),      
              state = "MD",
              survey = "acs1",
              year = 2022) %>% 
  mutate(year = 2022)
#Why does this only show 12 places?
# Per Derek: "Using 2022 1-year values will only get you the largest places in the state. Probably need to dial it back to an earlier year"

#GEOID will be key to matching the places up together. 

usa_2022 <- get_acs(geography = "state",
                variables = c(population = "B01003_001"),
                survey = "acs1",
                year = 2022) %>% 
  mutate(year = 2022)

#Notes
#I think I need to join the populations to the state_counts dataframe to then do a mutate and identify number of WBB players per capita in each state.  
# Would I need to change the city names to be accurate with the census stuff


```

Using 2021 data for Maryland and the 50 states.

```{r}

md_2021 <- get_acs(geography = "place",
                variables = c(population = "B01003_001"),      
              state = "MD",
              survey = "acs5",
              year = 2021) %>% 
  mutate(year = 2021)
#Why does this only show 12 places?
# Per Derek: "Using 2022 1-year values will only get you the largest places in the state. Probably need to dial it back to an earlier year"

#GEOID will be key to matching the places up together. 

usa_2021 <- get_acs(geography = "state",
                variables = c(population = "B01003_001"),
                survey = "acs5",
                year = 2021) %>% 
  mutate(year = 2021)

#Explain that these are 5-year rolling averages in the story


```

Next Steps
- Join the Census data to the appropreate dataframes
- Run code to genearte per capita numbers
- Write sentences to see what the data tells me
- Think of how I want to visualize the story and what my lede, body and structure will look like


