---
title: "wbb_roster_story"
author: "Dylan Manfre"
date: "2023-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

new section 

```{r}
# Libraries
library(tidyverse)
library(tidycensus)
library(janitor)
library(lubridate)
library(ipumsr)
library(usdata)
library(postmastr)
```

```{r}
# Load data for the 2022-23 season
wbb_rosters_22_23 <- read_csv("data/wbb_rosters_2022_23.csv")

```


# Organized Section
# Working Lede
# With the women's basketball season approaching, let's analyze player origins.
# Maryland has a diverse group of players this season.

```{r}
# Division I state counts 2022-23
d1_state_counts_22_23 <- wbb_rosters_22_23 %>%
  filter(division == "I") %>%
  group_by(state_clean) %>%
  count() %>%
  rename(number_of_players = n) %>%
  arrange(desc(number_of_players))

```


```{r}
# International Players 2022-23
international_players <- wbb_rosters_22_23 %>%
  filter(country_clean != "USA")

countries <- international_players %>%
  group_by(country_clean) %>%
  count() %>%
  arrange(desc(n))
```

# Where are D-I players from in Maryland
# Maryland has 133 Division I women's basketball players, and they come from various locations.
# About 18% of them play for Power 5 institutions.
```{r}
power_five <- c("SEC", "Big Ten", "Pac-12", "Big 12", "ACC")

from_maryland_p5 <- wbb_rosters_22_23 %>%
  filter(state_clean == "MD", conference %in% power_five)

# Percentage of MD players at P5 institutions
md_p5_perc <- (nrow(from_maryland_p5) / nrow(wbb_rosters_22_23)) * 100

# Which Power 5 conference produces the most WBB players?
conference_md <- from_maryland_p5 %>%
  group_by(conference) %>%
  count() %>%
  rename(players_from_md = n)

#Showing Division One from Maryland
division_one_md <- wbb_rosters_22_23 %>% 
  filter(state_clean == "MD", division == "I")

```

Showing number of power 5 vs Mid-Major
```{r}
p5_vs_mm <- division_one_md %>%
  mutate(p5_or_mm = case_when(
    conference %in% power_five ~ "power_five",
    TRUE ~ "mid-major"
  ))
```

Exploration: There are 109 players in Mid Major Programs from Maryland. 
```{r}
p5_vs_mm %>%  
  count(p5_or_mm)

```

cleaning high schools

What I'm doing here is standardizing the HS's of Maryland players 
```{r}

division_one_md_with_cleaned_high_school <- division_one_md %>%  
  mutate(hs_clean = case_when(
    hs_clean == "St. Frances Academy (Maryland)" ~ "St. Francis Academy",
    hs_clean == "Bishop McNamara High School" ~ "Bishop McNamara",
    hs_clean == "Bishop McNamara HS" ~ "Bishop McNamara",
    hs_clean == "Baltimore Polytechnic Institute" ~ "Baltimore Poly Institute",
    hs_clean == "Eleanor Roosevelt H.S." ~ "Eleanor Roosevelt",
    hs_clean == "Frederick High School" ~ "Frederick HS",
    hs_clean == "McDonogh High School" | hs_clean ==  "McDonogh School" ~ "McDonogh",
    hs_clean == "Paint Branch High School|Paintbranch" ~ "Paint Branch HS",
    hs_clean == "Riverdale Baptist School" ~ "Riverdale Baptist",
    hs_clean == "Roland Park County School" ~ "Roland Park Country School",
    hs_clean == "Saint Frances Academy" | hs_clean == "St. Frances Academy" | hs_clean == "St. Francis Academy" ~ "St. Frances Academy",
    hs_clean == "St. John's College" | hs_clean == "St. John's College High School" ~ "St. John's College HS",
    hs_clean == "Western High School" ~ "Western HS",
    TRUE ~ hs_clean
  ))

```


```{r}
# Which high school has the most players? Is there a pipeline between any schools?
#These highschools are mostly from Maryland but can also be in DC because players from Maryland 

high_schools_22_23 <- division_one_md_with_cleaned_high_school %>%
  group_by(hs_clean) %>%
  count() %>% 
  arrange(desc(n))

write_csv(high_schools_22_23, "data/high_schools_22_23.csv")

```
#New Hope Acadamy has the most wtih 6 players. Bishop McNamara has 4 players 

```{r}
#Showing top 10 Maryland towns from hometown 
view(md_d1_hometown)
```


```{r}
# All divisions with Maryland players
all_divison_md_players <- wbb_rosters_22_23 %>%
  filter(state_clean == "MD") %>%
  select(team, name, year_clean, position_clean, height_clean, hometown_clean, hs_clean, division, conference)
```

section on 2023-2024 season.

# Loading data for the 2023-2024 season
wbb_rosters_23_24 <- read_csv("data/wbb_rosters_2023_24.csv")

leaning the data based on what is in the wbb_1129 rmd.file

```{r}
wbb_rosters_23_24 <- wbb_rosters_23_24 %>% 
  mutate(hometown_cleaned = str_to_upper(str_replace(hometown, "\\.",""))) %>% 
  mutate(hometown_cleaned = str_replace(hometown_cleaned,"\\.","")) %>%
  mutate(hometown_cleaned = str_replace(hometown_cleaned,"/.*","")) %>%
  mutate(hometown_cleaned = str_replace(hometown_cleaned,"\\.$","")) %>%
  mutate(hometown_cleaned = str_replace(hometown_cleaned,",",", ")) %>%
  mutate(hometown_cleaned = str_replace(hometown_cleaned,"-"," ")) %>% 
  mutate(hometown_cleaned = str_squish(hometown_cleaned))

states_non_standard <- pm_append(type = "state", input = c("SD.", "MASS", "CALIF", "MICH", "NEB", "IND", "MINN", "ORE", "OHIO", "FLA", "MISS", "TENN", "ARIZ", "KAN", "ALA", "OKLA", "WIS", "ILL", " WASH", "ARK", "COLO", "NEV", "CONN", "WISC", "WVA", "DEL", "WYO", "CALI", "LOUIS", "VIRG", "MONT", "PENN", "TEX", "KANS", "NEBR", "IDA", "COL"), output = c("SD", "MA", "CA", "MI", "NE", "IN", "MN", "OR", "OH", "FL", "MS", "TN", "AZ", "KS", "AL", "OK", "WI", "IL", "WA", "AR", "CO", "NV", "CT", "WI", "WV", "DE", "WY", "CA", "LA", "VA", "MT", "PA", "TX", "KS", "NE", "ID", "CO"), locale = "us")

dict <- pm_dictionary(type='state', case = c("title", "upper", "lower"), append = states_non_standard)

wbb_rosters_23_24 <- wbb_rosters_23_24 %>% pm_identify(var="hometown_cleaned")

parsed <- wbb_rosters_23_24 %>% 
  pm_identify(var="hometown_cleaned") %>% 
  pm_prep(var="hometown_cleaned", type="street") %>% 
  pm_state_parse(dict)

wbb_rosters_23_24 <- left_join(wbb_rosters_23_24, parsed, by="pm.uid")

wbb_rosters_23_24 %>%
  filter(is.na(hometown_cleaned))

wbb_rosters_23_24 <- wbb_rosters_23_24 %>% separate(hometown, c('hometown', 'homestate'), sep=',', extra='merge')

wbb_rosters_23_24 <- wbb_rosters_23_24 %>% mutate(homestate=str_trim(homestate), homestate=str_replace(homestate, '\\.', '')) %>%
  mutate(homestate = case_when(is.na(abbr2state(homestate)) ~ homestate, TRUE ~ (abbr2state(homestate))))

wbb_rosters_23_24 <- wbb_rosters_23_24 %>% 
rename(cleaned_city = pm.address,
        cleaned_state = pm.state)

view(wbb_rosters_23_24)
view(wbb_rosters_22_23)
```

```{r}
# Merging division column from 22-23 to 23-24 dataframe
wbb_rosters_23_24 <- inner_join(wbb_rosters_23_24, divisions_and_teams, by = "team")
```

```{r}
# D1 state counts for 23-24 season
d1_state_counts_23_24 <- wbb_rosters_23_24 %>%
  filter(division == "I") %>%
  group_by(cleaned_state) %>%
  count() %>%
  rename(number_of_players = n) %>%
  arrange(desc(number_of_players))

```

```{r}
# D1 players from Maryland for 23-24 season
division_one_md_23_24 <- wbb_rosters_23_24 %>%
  filter(cleaned_state == "MD", division == "I")
```

```{r}
# High schools with D1 players for 23-24 season
high_schools_23_24 <- division_one_md_23_24 %>%
  group_by(high_school) %>%
  count() %>%
  arrange(desc(n))
```

```{r}

```


```{r}
# Exporting to CSV
write_csv(md_d1_hometown, "data/md_d1_hometown.csv")
write_csv(division_one_md, "data/division_one_md.csv")
write_csv(high_schools_22_23, "data/high_schools_22_23.csv")
write_csv(md_d1_hometown, "data/md_d1_hometown.csv")
write_csv(teams_md_players_22_23, "data/teams_md_players_22_23.csv")
```
  
